---
title: "TEMPEST2023_Fe"
author: "Ashleigh Montgomery"
date: "2024-06-30"
output: html_document
---

#SET UP
```{r}
#Set working directory
setwd("~/R Code/TEMPEST/Porewater/Fe")

library(googlesheets4)
library(googledrive)
library(tidyverse)
library(data.table)
library(plater)
library(readxl)
library(lubridate)
```

##Helpful functions
```{r}
#Function to calculate slope, intercept, and R-squared and return in data frame
#Fe 2 
cal_curve<-function(Conc_uM,ABS){
    model<-lm(ABS~Conc_uM)
    slope<-coef(model)[[2]]
    intercept<-coef(model)[[1]]
     R2<-summary(model)$adj.r.squared
  return(data.frame(slope=slope,intercept=intercept,R2=R2))
}


#function to calculate percent difference
calc_percent_diff<-function(dup,original){
  avg<-(dup+original)/2
  percent_diff<-((abs(original-dup))/avg)*100
  return(percent_diff=percent_diff)
}

#function to calculate percent error
calc_percent_error<-function(chk,expected){
  percent_error<-(abs((chk-expected))/expected)*100
  return(percent_error=percent_error)
}
```



##Needed values
```{r}
#set absorbance threshold (if surpassed, use high standard curve)
#May need to be adjusted between runs 
Fe2_threshold<-0.2 
Fe3_threshold<-0.2
```



#UPLOAD DATA
##Sample sheet and inventory
```{r}
#google sheet file with porewater inventory
#needed to get sample volumes to appropriately calculate dilutions

#Link to inventory on google drive
inventory_directory <- "https://docs.google.com/spreadsheets/d/1sFWq-WKhemPzbOFInqhCu_Lx0lsO6a_Z/edit#gid=496164093"

#downloads current version of file and creates xlsx file in working directory
drive_download(inventory_directory, path="./porewaterinventory.xlsx", overwrite = TRUE)

#saves sheet of interest, filters for Fe 
inventory<-read_excel("porewaterinventory.xlsx",sheet = "Individual",skip = 3)%>%
  select(Sample_ID, Volume_mL,  Evacuation_date_YYYMMDD,Collection_Date_YYYYMMDD, Collection_End_Time_24hrs, EST_EDT)%>%
  filter(str_detect(Sample_ID, "FE")) %>%
  rename(sample_name = Sample_ID,
         evacuation_date = Evacuation_date_YYYMMDD,
         collection_date = Collection_Date_YYYYMMDD,
         time = Collection_End_Time_24hrs,
         tz = EST_EDT) %>%
  mutate(evacuation_date = lubridate::as_date(as.character(evacuation_date), format = "%Y%m%d"),
         collection_date = lubridate::as_date(as.character(collection_date), format = "%Y%m%d"),
         elapsed_time = lubridate::days(collection_date - evacuation_date)) %>%
  select(sample_name, evacuation_date, collection_date, time, tz, elapsed_time,Volume_mL)
head(inventory)
```

##Fe plates
```{r}
#Standards
std_plates<-read_plates(files = c("TEMPEST_FE_20230607_Stds1.csv",
                            "TEMPEST_FE_20230530_Stds1.csv",
                            "TEMPEST_FE_20230606_Stds1.csv",
                            "TEMPEST_FE_20230608_Stds.csv",
                            "TEMPEST_FE_20230609_Stds1.csv",
                            "TEMPEST_FE_20230808_Stds1.csv",
                            "TEMPEST_FE_20230908_Stds1.csv",
                            "TEMPEST_FE_20231208_Stds1.csv",
                            "TEMPEST_FE_20240510_Stds1.csv"),
                  plate_names = c("20230607",
                                  "20230530",
                                  "20230606",
                                  "20230608",
                                  "20230609",
                                  "20230808",
                                  "20230908",
                                  "20231208",
                                  "20240510"),
                  well_ids_column = "Wells",
                  sep=",")

head(std_plates)

sample_plates<-read_plates(files = c("TEMPEST_FE_20230607_Samples1.csv",
                               "TEMPEST_FE_20230607_Samples2.csv",
                            "TEMPEST_FE_20230530_Samples1.csv",
                            "TEMPEST_FE_20230606_Samples1.csv",
                            "TEMPEST_FE_20230608_Samples1.csv",
                            "TEMPEST_FE_20230608_Samples2.csv",
                            "TEMPEST_FE_20230609_Samples1.csv",
                            "TEMPEST_FE_20230808_Samples1.csv",
                            "TEMPEST_FE_20230808_Samples2.csv",
                            "TEMPEST_FE_20230808_Samples3.csv",
                            "TEMPEST_FE_20230808_Samples4.csv",
                            "TEMPEST_FE_20230908_Samples1.csv",
                            "TEMPEST_FE_20230908_Samples2.csv",
                            "TEMPEST_FE_20231208_Samples1.csv",
                            "TEMPEST_FE_20240510_Samples1.csv"),
                  plate_names = c("20230607",
                                  "20230607",
                                  "20230530",
                                  "20230606",
                                  "20230608",
                                  "20230608",
                                  "20230609",
                                  "20230808",
                                  "20230808",
                                  "20230808",
                                  "20230808",
                                  "20230908",
                                  "20230908",
                                  "20231208",
                                  "20240510"),
                  well_ids_column = "Wells",
                  sep=",")%>%
  select(!Conc_uM)%>%
  mutate(Fe_2_Curve=if_else(Abs1>Fe2_threshold,"High","Low"), #flags cal curve
         Fe_3_Curve=if_else(Abs2>Fe3_threshold,"High","Low"))
  
head(sample_plates)
```

#PLOT STANDARDS
##FE 2
```{r}
#plot curve
std_plates%>%
  filter(str_detect(std_plates$IDs,"Fe_2_|HCl"))%>% #Filters out Fe 2 standards
  mutate(Curve=ifelse(Abs1>Fe2_threshold,"High","Low"))%>% #separates into low and high curves
  na.omit()%>%
  ggplot(aes(Conc_uM,Abs1))+
  facet_grid(Curve~Plate)+
  geom_point()+
  geom_smooth(method="lm",se=FALSE)+
  ggtitle("Fe 2 Standards")

#calculate R2, slope, and intercept
Fe2_stds<-std_plates%>%
  filter(str_detect(std_plates$IDs,"Fe_2_|HCl"))%>%
  mutate(Curve=ifelse(Abs1>Fe2_threshold,"High","Low"))%>%
  na.omit()%>%
  group_by(Plate,Curve)%>%
  mutate(cal_curve(Conc_uM,Abs1))%>%
  filter(row_number()==1)%>%
  select(Plate,Curve,slope,intercept,R2)%>%
  rename(Fe_2_Curve=Curve,Fe_2_slope=slope,Fe_2_intercept=intercept,Fe_2_R2=R2)

Fe2_stds

```

##FE 3 
```{r}
#plot curve
std_plates%>%
  filter(str_detect(std_plates$IDs,"Fe_3_|HCl"))%>% #Filters out Fe 3 standards
  mutate(Curve=ifelse(Abs2>Fe3_threshold,"High","Low"))%>% #separates into low and high curves
  na.omit()%>%
  ggplot(aes(Conc_uM,Abs2))+
  facet_grid(Curve~Plate)+
  geom_point()+
  geom_smooth(method="lm",se=FALSE)+
  ggtitle("Fe 3 standards")

#calculate R2, slope, and intercept
Fe3_stds<-std_plates%>%
  filter(str_detect(std_plates$IDs,"Fe_3_|HCl"))%>%
  mutate(Curve=ifelse(Abs2>Fe3_threshold,"High","Low"))%>%
  na.omit()%>%
  group_by(Plate,Curve)%>%
  mutate(cal_curve(Conc_uM,Abs2))%>%
  filter(row_number()==1)%>%
  select(Plate,Curve,slope,intercept,R2)%>%
  rename(Fe_3_Curve=Curve,Fe_3_slope=slope,Fe_3_intercept=intercept,Fe_3_R2=R2)

Fe3_stds #highs look terrible. Anything that falls into this range should be diluted and rerun 

```

#Calculations of sample concentrations 
```{r}
#add calibration curve information and convert abs to concentrations (uM)

head(Fe2_stds) #Fe 2 standard R2, intercept, and slope data frame
head(Fe3_stds) #Fe 3 standard R2, intercept, and slope data frame
head(sample_plates) #samples

dat<-left_join(sample_plates,Fe2_stds,by=c("Plate","Fe_2_Curve"))%>% #add cal curve info
  left_join(Fe3_stds,by=c("Plate","Fe_3_Curve"))%>%
  mutate(Fe_2=((Abs1-Fe_2_intercept)/Fe_2_slope)*Dilutions, #calculate concentrations
         Fe_Total=((Abs2-Fe_3_intercept)/Fe_3_slope)*Dilutions,
         Fe_3=Fe_Total-Fe_2)%>%
  mutate(Fe_2=if_else(Fe_2<0,0,Fe_2), #change negative values to zero
         Fe_Total=if_else(Fe_Total<0,0,Fe_Total),
         Fe_3=if_else(Fe_3<0,0,Fe_3))%>%
  mutate(Fe_2_info=if_else(Fe_2<0,"bdl","within range"), #add info about sample
         Fe_3_info=if_else(Fe_3<0,"bdl","within range"))%>%
  filter(!is.na(IDs), #remove NAs
         !is.na(Abs1),
         !is.na(Abs2))
head(dat)
```

```{r}
#Average concentrations across wells and calculate standard deviation
dat.mean<-dat%>%
  group_by(Plate,IDs)%>%
  summarise(Fe_2_uM=mean(Fe_2),
            Fe_2_sd=sd(Fe_2),
            Fe_Total_uM=mean(Fe_Total),
            Fe_Total_sd=sd(Fe_Total),
            Fe_3_uM=mean(Fe_3),
            Fe_3_sd=sd(Fe_3),
            Abs1=mean(Abs1),
            Abs2=mean(Abs2))%>%
  mutate(Fe_3_info=case_when(
    Abs2 > Fe3_threshold ~ "Out of Range",
    Abs2 < Fe3_threshold ~ "In Range"))

head(dat.mean)

#samples with high standard deviation
dat.mean%>%
  filter(Fe_2_sd>5|Fe_3_sd>5)
```


#combine samples with inventory info, important to correct for dilution by HCl due to different volumes of sample collected
```{r}
hcl.vol<-0.075 #mL of HCl in prepped vial

head(inventory)
#separate into event and non-event samples

event_inventory<-inventory%>%
  filter(str_detect(sample_name,"_2023"))%>%
  filter(str_detect(sample_name,"PM|AM"))%>%
  separate_wider_delim(sample_name,delim="_",names=c("Project","Treatment","Sample Type","Grid","Depth","Analyte","Collection_Date","Collection_Time"))

nonevent_inventory<-inventory%>%
   filter(str_detect(sample_name,"_2023"))%>%
  filter(!str_detect(sample_name,"PM|AM"))%>%
  separate_wider_delim(sample_name,delim="_",names=c("Project","Treatment","Sample Type","Grid","Depth","Analyte","Collection_Date"))%>%
  mutate(Collection_Time="1200")

#put back together
Tempest2.0_inventory<-rbind(event_inventory,nonevent_inventory)


head(dat.mean)
#separate into event and non-event samples
event_samples<-dat.mean%>%
  filter(!str_detect(IDs,"Fe|HCl|ESTUARY|SOURCE|spk|dup"))%>%
  filter(str_detect(IDs,"_0800|1600"))%>%
  separate_wider_delim(IDs,delim = "_",names = c("Treatment","Grid","Collection_Date","Collection_Time"))%>%
  mutate(Collection_Time=if_else(Collection_Time=="0800","AM","PM"))

nonevent_samples<-dat.mean%>%
  filter(!str_detect(IDs,"Fe|HCl|ESTUARY|SOURCE|spk|dup|_0800|1600|AQUIFER"))%>%
  separate_wider_delim(IDs,delim = "_",names = c("Treatment","Grid","Collection_Date"))%>%
  mutate(Collection_Time="1200")

TEMPEST2.0_samples<-rbind(event_samples,nonevent_samples)

#combine inventory to sample data 
dat.final<-left_join(Tempest2.0_inventory,TEMPEST2.0_samples,by=c("Treatment","Grid","Collection_Date","Collection_Time"))%>%
  mutate(DF=(Volume_mL+hcl.vol)/(Volume_mL),
         Fe_2_uM=Fe_2_uM*DF,
         Fe_Total_uM=Fe_Total_uM*DF,
         Fe_3_uM=Fe_3_uM*DF)%>%
  mutate(IDs=paste(Treatment,Grid,Collection_Date,sep = "_"))%>%
    mutate(time=format(strptime(substr(as.POSIXct(sprintf("%04.0f",time),format="%H%M"),12,16),"%H:%M"),"%I:%M %p"))%>%
  mutate(DateTime=paste(collection_date,time))

head(dat.final)

```

```{r}
#quick first look at data (just samples)
Fe2<-dat.final%>%
  filter(!Plate=="NA")%>%
  ggplot(aes(x=IDs,y=Fe_2_uM))+
  geom_point(aes(group=Plate),size=4)+
  facet_wrap(~Plate,scales = "free_x")+
  labs(y="Fe (II) uM",x="Sample ID")+
  geom_errorbar(aes(ymin=Fe_2_uM-Fe_2_sd,
                    ymax=Fe_2_uM+Fe_2_sd),
                width=0.3,
                position=position_dodge(.1))+
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1))+
  ggtitle("Fe 2")

Fe2

Fe3<-dat.final%>%
  filter(!Plate=="NA")%>%
  ggplot(aes(x=IDs,y=Fe_3_uM))+
  geom_point(aes(group=Plate),size=4)+
  facet_wrap(~Plate,scales="free_x")+
  labs(y="Fe (III) uM",x="Sample ID")+
  geom_errorbar(aes(ymin=Fe_3_uM-Fe_3_sd,
                    ymax=Fe_3_uM+Fe_3_sd),
                width=0.3,
                position=position_dodge(.1))+
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1))+
  ggtitle("Fe 3")

Fe3

```

```{r}
#export data 
write.csv(dat.final,file="TEMPEST2023_Fe.csv")
```



#QA/QC
#Reduction Efficiency
```{r}
#pull out 60 uM Fe II and Fe III standards
Fe_2_60uM<-std_plates%>%
  filter(str_detect(std_plates$IDs,"Fe_2"))%>%
  filter(Conc_uM==60)%>%
  na.omit()%>%
  group_by(Plate)%>%
  summarise(Abs1=mean(Abs1),Abs2=mean(Abs2))

Fe_2_60uM

Fe_3_60uM<-std_plates%>%
  filter(str_detect(std_plates$IDs,"Fe_3"))%>%
  filter(Conc_uM==60)%>%
  na.omit()%>%
  group_by(Plate)%>%
  summarise(Abs1=mean(Abs1),Abs2=mean(Abs2))

Fe_3_60uM
```




```{r}
#calculate reduction effeciency
red_eff<-data.frame(matrix(ncol = 0, nrow = nrow(Fe_3_60uM)))

red_eff$Plate <- Fe_2_60uM$Plate
red_eff$FE2 <- Fe_2_60uM$Abs1
red_eff$FE3 <- Fe_3_60uM$Abs2

Fe_red_eff<-red_eff%>%
  group_by(Plate)%>%
  mutate(Eff=(FE3/FE2)*100, 
         Eff2=((FE3/0.8)/FE2)*100)%>% ## accounts for dilution by reagents
  mutate(Red_eff=case_when(
    Eff < 80 | Eff > 120 ~ "BAD",
    Eff > 80 & Eff < 90 ~ "OK",
    Eff <= 120 & Eff > 110 ~ "OK",
    Eff >= 90 & Eff <=110 ~ "GOOD"
  ))

Fe_red_eff
```



#Duplicates
```{r}
#pull out dups
dups<-dat.mean%>%
  filter(str_detect(IDs,"_dup"))
head(dups)

#remove these from dat.mean
dat2<-dat.mean%>%
  filter(!str_detect(IDs,"_dup|_spk"))
head(dat2)


#remove the dup from these IDs so we will have duplicates 
dups$IDs<-gsub("_dup","",as.character(dups$IDs))
dups <- dups[ ,c(1,2,3,7)]
colnames(dups) <- c("Plate",'IDs', 'Fe_2_mean_dup', "Fe_3_mean_dup")
head(dups)

#add dups back to data set
QAdup<-merge(dat2,dups)
head(QAdup)

#calculate percent difference between dups
QAdup<-QAdup%>%
  mutate(Fe_2_pdiff=calc_percent_diff(Fe_2_mean_dup,Fe_2_uM),
            Fe_3_pdiff=calc_percent_diff(Fe_3_mean_dup,Fe_3_uM))%>%
  #ADD QAQC FLAGS
  mutate(dup2=case_when(
    Fe_2_pdiff <= 10 ~ "GOOD",
    Fe_2_pdiff >10 & Fe_2_pdiff <=15 ~ "OK",
    Fe_2_pdiff >15 ~ "BAD"
  ),
  dup3=case_when(
    Fe_3_pdiff <= 10 ~ "GOOD",
    Fe_3_pdiff >10 & Fe_2_pdiff <=15 ~ "OK",
    Fe_3_pdiff >15 ~ "BAD",
  ))%>%
  select(Plate,IDs,Fe_2_uM,Fe_2_mean_dup,Fe_2_pdiff,dup2,
         Fe_3_uM,Fe_3_mean_dup,Fe_3_pdiff,dup3)
QAdup
```


#Spikes
```{r}
#calculate DF of spiked sample
sample_vol<-160 #uL
spike_vol<-55 #uL
DF<-(sample_vol+spike_vol)/sample_vol

#spike concentration
spk_uM<-60

#pull out spikes
spks<-dat.mean%>%
  filter(str_detect(IDs,"_spk"))%>%
  mutate(Fe_2_uM=Fe_2_uM*DF)
head(spks)

#remove spikes from dat.mean
dat2<-dat.mean%>%
  filter(!str_detect(IDs,"_spk"),
         !str_detect(IDs,"_dup"))
head(dat2)

#remove _spk from names 
spks$IDs<-gsub("_spk","",as.character(spks$IDs))
spks<-spks[ ,c(1,2,3)]
colnames(spks)<-c("Plate","IDs","Fe_2_mean_spk")

#combine spks with original data set 
QAspks<-merge(dat2,spks)
head(QAspks)

#Calculate % recovery
QAspks<-QAspks%>%
  mutate(percent_recovery=((Fe_2_mean_spk-Fe_2_uM)/spk_uM)*100)%>%
  mutate(spks=case_when(
    percent_recovery < 80 | percent_recovery > 120 ~ "BAD",
    percent_recovery > 80 & percent_recovery < 90 ~ "OK",
    percent_recovery <= 120 & percent_recovery > 110 ~ "OK",
    percent_recovery >= 90 & percent_recovery <=110 ~ "GOOD"
  ))

head(QAspks)


```


#check standards 
```{r}
#calculate standard concentrations from cal curve
stds<-std_plates%>%
  mutate(Fe_2_Curve=if_else(Abs1>Fe2_threshold,"High","Low"), #flags cal curve
         Fe_3_Curve=if_else(Abs2>Fe3_threshold,"High","Low"))%>%
  left_join(Fe2_stds,by=c("Plate","Fe_2_Curve"))%>%
  left_join(Fe3_stds,by=c("Plate","Fe_3_Curve"))%>%
  mutate(Fe_2_uM_std=ifelse(str_detect(IDs,"_2_|HCl"),
         ((Abs1-Fe_2_intercept)/Fe_2_slope)*Dilutions,
         NA),
         Fe_3_uM_std=ifelse(str_detect(IDs,"_3_|HCl"),
                                       ((Abs2-Fe_3_intercept)/Fe_3_slope)*Dilutions,
                                       NA))%>%
  group_by(Plate,IDs)%>%
  summarise(Fe_2_uM_std=mean(Fe_2_uM_std),
            Fe_3_uM_std=mean(Fe_3_uM_std))

head(stds)

```




```{r}
#pull out check standards
chk<-dat.mean%>%
  filter(str_detect(IDs,"uM|HCl"))%>%
  mutate(Concentration=case_when(
    str_detect(IDs,"0.6") ~ 0.6,
    str_detect(IDs,"250") ~ 250,
    str_detect(IDs,"3 uM") ~ 3,
    str_detect(IDs,"60") ~ 60,
    str_detect(IDs,"18") ~ 18,
    str_detect(IDs,"9") ~ 9,
    str_detect(IDs,"HCl") ~ 0
  ))%>%
  select(Plate,IDs,Fe_2_uM,Fe_3_uM,Concentration)

head(chk)
```

```{r}
#put together with standards
QAchks<-left_join(stds,chk,by=c("Plate","IDs"))
head(QAchks)

#calculate % difference
QAchks<-QAchks%>%
  filter(!str_detect(IDs,"HCl"))%>%
  mutate(pdiff=if_else(str_detect(IDs,"Fe_2_"),
                       calc_percent_diff(Fe_2_uM,Fe_2_uM_std),
                       calc_percent_diff(Fe_3_uM,Fe_3_uM_std)),
         perror=if_else(str_detect(IDs,"Fe_2_"),
                        calc_percent_error(Fe_2_uM,Concentration),
                        calc_percent_error(Fe_3_uM,Concentration)))%>%
  mutate(QA_pdiff=case_when(
        pdiff <= 10 ~ "GOOD",
    pdiff >10 & pdiff <=15 ~ "OK",
    pdiff >15 ~ "BAD"
  ),
  QA_perror=case_when(
    perror <= 10 ~ "GOOD",
    perror>10 & perror <=15 ~ "OK",
    perror >15 ~ "BAD"
  ))%>%
  drop_na(QA_pdiff)

head(QAchks)
```

*Figure out a way to summarize QA/QC in one table






